PANCHANGE_descriptives

### 2020-09-15 PANCHANGE_exclusion_descriptives.Rmd
test run of all scripts

## Participants per Sample
Eating Disorders Genetics Initative
Genetic Links to Anxiety and Depression
National BioResource
Rapid Assessment of MP
```{r sample frequencies}
dat %>%
  freq(Sample)
```

## Age categories

Age bar plot for whole sample (not grouped)
Summary dataset for age and display it as a bar plot
```{r age_category prop total (ungrouped)}
age_category_prop_total <- dat %>%
  count(age_category) %>%
  mutate(Prop = round(n/sum(n), 2))

```{r age_category per Sample chisquare test}
age_category_per_Sample_chissquare_test <- dat %>%
  count(age_category, Sample) %>%
  spread(key = Sample, value = n) %>%
  select(-age_category) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  ) %>%
  chisq.test() %>%
  glance()

```{r Ethnicity_collapsed per Sample chisquare test}
Ethnicity_collapsed_per_Sample_chissquare_test <- dat %>%
  count(Ethnicity_collapsed, Sample) %>%
  spread(key = Sample, value = n) %>%
  select(-Ethnicity_collapsed) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  ) %>%
  chisq.test() %>%
  glance()

### pie charts

```{r Ethnicity_collapsed prop per total summary}

Ethnicity_collapsed_prop_Sample <- dat %>%
  count(Ethnicity_collapsed) %>%
  drop_na() %>%
  mutate(freq = round(n/sum(n), 2))

kable(Ethnicity_collapsed_prop_Sample)
Ethnicity_collapsed_prop_Sample

```{r ethnicity pie chart biggest group label}

eth_pie_alllabel <-

  Ethnicity_collapsed_prop_Sample %>%
   mutate(Ethnicity_collapsed = fct_reorder(Ethnicity_collapsed,-n)) %>%

eth_pie_alllabel

```{r ethnicity pie chart no label}

eth_pie <-

    Ethnicity_collapsed_prop_Sample %>%
   mutate(Ethnicity_collapsed = fct_reorder(Ethnicity_collapsed,-n)) %>%

eth_pie

```{r Gender_collapsed per Sample chisquare test}
Gender_collapsed_per_Sample_chissquare_test <- dat %>%
  count(Gender_collapsed, Sample) %>%
  spread(key = Sample, value = n) %>%
  select(-Gender_collapsed) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  ) %>%
  chisq.test() %>%
  glance()

### pie charts

```{r Gender_collapsed prop per total}

Gender_collapsed_prop_total <- dat %>%
  count(Gender_collapsed) %>%
  mutate(freq = round(n/sum(n), 2))

kable(Gender_collapsed_prop_total)
Gender_collapsed_prop_total

```{r gender pie chart - all labels}

gen_pie_alllabel <- 
  ggplot(Gender_collapsed_prop_total, aes(x="", y=freq, fill=Gender_collapsed)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette5) +
  scale_color_manual(values =  COPINGpalette5) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported gender",
                      sep=" "),
        subtitle = paste("n", sum(Gender_collapsed_prop_total$n),
                         sep = " "),
        fill = "Self reported gender")  +
    geom_text(aes(y = c(0.85,0.35,0),
                  x = c(1.2,1.2,1.8),
                  label= c(paste0(round((freq[1:2]*100),0),"%"),paste0(round((freq[3]*100),1),"%"))),
             size=5,
             face="bold"
             )

gen_pie_alllabel

pie chart - only labelling two largest groups (male and female)

```{r gender pie chart - male/female labels only}

gen_pie <- 
  ggplot(Gender_collapsed_prop_total, aes(x="", y=freq, fill=Gender_collapsed)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette5) +
  scale_color_manual(values =  COPINGpalette5) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported gender",
                      sep=" "),
        subtitle = paste("n", sum(Gender_collapsed_prop_total$n),
                         sep = " "),
        fill = "Self reported gender")  +
    geom_text(aes(y = c(0.85,0.35,0),
                  x = c(1.2,1.2,1.8),
                  label= c(paste0(round((freq[1:2]*100),0),"%"),"")),
             size=5,
             face="bold"
             )

gen_pie

```{r gender pie chart - no labels }

gen_pie_nolabel <- 
  ggplot(Gender_collapsed_prop_total, aes(x="", y=freq, fill=Gender_collapsed)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette5) +
  scale_color_manual(values =  COPINGpalette5) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported gender",
                      sep=" "),
        subtitle = paste("n", sum(Gender_collapsed_prop_total$n),
                         sep = " "),
        fill = "Self reported gender")

gen_pie_nolabel

```{r highest_education per Sample chisquare test}
highest_education_per_Sample_chissquare_test <- dat %>%
  count(highest_education, Sample) %>%
  spread(key = Sample, value = n) %>%
  select(-highest_education) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  ) %>%
  chisq.test() %>%
  glance()

```{r highest_education by Sample count plot}
highest_education_by_Sample_count_plot <- highest_education_per_Sample_prop_Sample %>%
  ggplot(mapping = aes(
    x = highest_education,
    y = n)
  ) +
  geom_bar(
    aes(fill = Sample),
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "highest_education",
    y = "Number of participants",
    fill = "Sample") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette6
  ) +
  coord_flip()

### pie chart

```{r highest education prop per total}

highest_education_prop_total <- dat %>%
  count(highest_education) %>%
  mutate(freq = round(n/sum(n), 2)) %>% 
  drop_na()

```{r binary education pie chart all labels}

edu_pie_binlabel <-

    edubindat %>%
    mutate(demographics.education_binary = fct_reorder(demographics.education_binary,-n)) %>%

  ggplot(aes(x="", y=freq, fill=demographics.education_binary)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette7) +
  scale_color_manual(values =  COPINGpalette7) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported highest level of education, above or below bachelor degree or equivalent",
                      sep=" "),
        subtitle = paste("n", sum(edubindat$n),
                         sep = " "),
        fill = "Self reported level of education")  +
    geom_text(aes(y = c(0.25,0.75),
                  label= paste0(round((freq*100),0),"%")),
             size=5,
             face="bold"
             )

edu_pie_binlabel

```{r education pie chart no labels }

edu_pie_nolabel <-

  ggplot(aes(x="", y=freq, fill=demographics.highest_level_of_education)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette7) +
  scale_color_manual(values =  COPINGpalette7) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported level of education",
                      sep=" "),
        subtitle = paste("n", sum(highest_education_prop_total$n),
                         sep = " "),
        fill = "Self reported level of education")

```{r key_worker per Sample chisquare test}
key_worker_per_Sample_chissquare_test <- dat %>%
  count(key_worker, Sample) %>%
  spread(key = Sample, value = n) %>%
  select(-key_worker) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  ) %>%
  chisq.test() %>%
  glance()

```{r key_worker by Sample count plot}
key_worker_by_Sample_count_plot <- key_worker_per_Sample_prop_Sample %>%
  ggplot(mapping = aes(
    x = key_worker,
    y = n)
  ) +
  geom_bar(
    aes(fill = Sample),
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "key_worker",
    y = "Number of participants",
    fill = "Sample") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette6
  ) +
  coord_flip()

@@

```{r Disorder_inc by Ethnicity_collapsed plot}
Disorder_inc_by_Ethnicity %>%
  ggplot(mapping = aes(
    x = Disorder_inc,
    y = n)
  ) +
  geom_bar(
    aes(fill = Ethnicity_collapsed),
    #    position = "dodge",
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Disorders",
    y = "Number of participants",
    fill = "Ethnicity_collapsed") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette6,
    #    labels = c("Prepandemic", "Retrospective")
  ) +
  coord_flip()
```

>>>>

## Participants per Sample
Eating Disorders Genetics Initative
Genetic Links to Anxiety and Depression
National BioResource
Rapid Assessment of MP
```{r participants per subsample frequencies}
dat %>%
  freq(Sample)
```

@@

```{r Disorder_inc by Ethnicity_collapsed plot}
Disorder_inc_by_Ethnicity %>%
  ggplot(mapping = aes(
    x = Disorder_inc,
    y = n)
  ) +
  geom_bar(
    aes(fill = Ethnicity_collapsed),
    #    position = "dodge",
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Disorders",
    y = "Number of participants",
    fill = "Ethnicity_collapsed") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette6
  ) +
  coord_flip()
```

## Age categories
Age bar plot for whole sample (not grouped)
Summary dataset for age and display it as a bar plot
```{r age_category prop total (ungrouped)}
age_category_prop_total <- dat %>%
  count(age_category) %>%
  mutate(Prop = round(n/sum(n), 2))

## Education

### Pie charts

```{r highest education prop per total}
highest_education_prop_total <- dat %>%
  count(highest_education) %>%
  mutate(freq = round(n/sum(n), 2)) %>% 
  drop_na()

```{r binary education pie chart all labels}
edu_pie_binlabel <- 
    edubindat %>%
    mutate(demographics.education_binary = fct_reorder(demographics.education_binary,-n)) %>%
  ggplot(aes(x="", y=freq, fill=demographics.education_binary)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette7) +
  scale_color_manual(values =  COPINGpalette7) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported highest level of education, above or below bachelor degree or equivalent",
                      sep=" "),
        subtitle = paste("n", sum(edubindat$n),
                         sep = " "),
        fill = "Self reported level of education")  +
    geom_text(aes(y = c(0.25,0.75),
                  label= paste0(round((freq*100),0),"%")),
             size=5,
             face="bold"
             )

edu_pie_binlabel
```

```{r education pie chart no labels }
edu_pie_nolabel <-

  ggplot(aes(x="", y=freq, fill=demographics.highest_level_of_education)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette7) +
  scale_color_manual(values =  COPINGpalette7) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported level of education",
                      sep=" "),
        subtitle = paste("n =", sum(highest_education_prop_total$n),
                         sep = " "),
        fill = "Self reported level of education")

edu_pie_nolabel
```

## Ethnicity_collapsed
### Pie charts

```{r Ethnicity_collapsed prop per total summary}
Ethnicity_collapsed_prop_Sample <- dat %>%
  count(Ethnicity_collapsed) %>%
  drop_na() %>%
  mutate(freq = round(n/sum(n), 2))

kable(Ethnicity_collapsed_prop_Sample)
Ethnicity_collapsed_prop_Sample
```

```{r ethnicity pie chart biggest group label}
eth_pie_alllabel <- 
  Ethnicity_collapsed_prop_Sample %>%
   mutate(Ethnicity_collapsed = fct_reorder(Ethnicity_collapsed,-n)) %>%

eth_pie_alllabel
```

Pie chart - no labels

```{r ethnicity pie chart no label}
eth_pie_nolabel <- 
  Ethnicity_collapsed_prop_Sample %>%
   mutate(Ethnicity_collapsed = fct_reorder(Ethnicity_collapsed,-n)) %>%

eth_pie_nolabel
```

## Gender collapsed
### Pie charts
```{r Gender_collapsed prop per total}
Gender_collapsed_prop_total <- dat %>%
  count(Gender_collapsed) %>%
  mutate(freq = round(n/sum(n), 2))

kable(Gender_collapsed_prop_total)
Gender_collapsed_prop_total
```

```{r gender pie chart - all labels}
gen_pie_alllabel <- 
  ggplot(Gender_collapsed_prop_total, aes(x="", y=freq, fill=Gender_collapsed)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette5) +
  scale_color_manual(values =  COPINGpalette5) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported gender",
                      sep=" "),
        subtitle = paste("n", sum(Gender_collapsed_prop_total$n),
                         sep = " "),
        fill = "Self reported gender")  +
    geom_text(aes(y = c(0.85,0.35,0),
                  x = c(1.2,1.2,1.8),
                  label= c(paste0(round((freq[1:2]*100),0),"%"),paste0(round((freq[3]*100),1),"%"))),
             size=5,
             face="bold"
             )

gen_pie_alllabel
```

Pie chart - only labelling two largest groups (male and female)
```{r gender pie chart - male/female labels only}
gen_pie <- 
  ggplot(Gender_collapsed_prop_total, aes(x="", y=freq, fill=Gender_collapsed)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette5) +
  scale_color_manual(values =  COPINGpalette5) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported gender",
                      sep=" "),
        subtitle = paste("n", sum(Gender_collapsed_prop_total$n),
                         sep = " "),
        fill = "Self reported gender")  +
    geom_text(aes(y = c(0.85,0.35,0),
                  x = c(1.2,1.2,1.8),
                  label= c(paste0(round((freq[1:2]*100),0),"%"),"")),
             size=5,
             face="bold"
             )

gen_pie
```

Pie chart -  no labels
```{r gender pie chart - no labels }
gen_pie_nolabel <- 
  ggplot(Gender_collapsed_prop_total, aes(x="", y=freq, fill=Gender_collapsed)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette5) +
  scale_color_manual(values =  COPINGpalette5) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported gender",
                      sep=" "),
        subtitle = paste("n", sum(Gender_collapsed_prop_total$n),
                         sep = " "),
        fill = "Self reported gender")

gen_pie_nolabel
```

## Chi square test
```{r age_category per Sample chisquare test}
age_category_per_Sample_chissquare_test <- dat %>%
  count(age_category, Sample) %>%
  spread(key = Sample, value = n) %>%
  select(-age_category) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  ) %>%
  chisq.test() %>%
  glance()

## Chi square test
```{r Ethnicity_collapsed per Sample chisquare test}
Ethnicity_collapsed_per_Sample_chissquare_test <- dat %>%
  count(Ethnicity_collapsed, Sample) %>%
  spread(key = Sample, value = n) %>%
  select(-Ethnicity_collapsed) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  ) %>%
  chisq.test() %>%
  glance()

## Chi square test
```{r Gender_collapsed per Sample chisquare test}
Gender_collapsed_per_Sample_chissquare_test <- dat %>%
  count(Gender_collapsed, Sample) %>%
  spread(key = Sample, value = n) %>%
  select(-Gender_collapsed) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  ) %>%
  chisq.test() %>%
  glance()

## Chi square test
```{r highest_education per Sample chisquare test}
highest_education_per_Sample_chissquare_test <- dat %>%
  count(highest_education, Sample) %>%
  spread(key = Sample, value = n) %>%
  select(-highest_education) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  ) %>%
  chisq.test() %>%
  glance()

## Bar plots
```{r highest_education by Sample count plot}
highest_education_by_Sample_count_plot <- highest_education_per_Sample_prop_Sample %>%
  ggplot(mapping = aes(
    x = highest_education,
    y = n)
  ) +
  geom_bar(
    aes(fill = Sample),
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "highest_education",
    y = "Number of participants",
    fill = "Sample") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette6
  ) +
  coord_flip()

## Chi square test
```{r key_worker per Sample chisquare test}
key_worker_per_Sample_chissquare_test <- dat %>%
  count(key_worker, Sample) %>%
  spread(key = Sample, value = n) %>%
  select(-key_worker) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  ) %>%
  chisq.test() %>%
  glance()

## Bar plots
```{r key_worker by Sample count plot}
key_worker_by_Sample_count_plot <- key_worker_per_Sample_prop_Sample %>%
  ggplot(mapping = aes(
    x = key_worker,
    y = n)
  ) +
  geom_bar(
    aes(fill = Sample),
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "key_worker",
    y = "Number of participants",
    fill = "Sample") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette6
  ) +
  coord_flip()

```{r Save four cohorts as rds file}
saveRDS(object = dat, file = paste0(data_path, "/four_cohorts_variables_exclusion.rds"))
```
